Embracing the Laws of Physics: \\ Three Reversible Models of Computation

### 2018-06-06 survey/survey.tex
one more pass over sec 1

\noindent These quotes by Toffoli and Feynman both highlight the
consequences of two obvious observations: (i) all the operations that
a computer performs reduce to basic physical operations; and (ii)
there is a mismatch between the logical operations of a typical model
of computation (which are logically irreversible) and the fundamental
laws of physics (which are reversible). One could certainly dismiss
the mismatch as irrelevant to the practice of computing but our thesis
is that the next computing revolution is likely to be founded on
revised models of computation that are designed to in closer harmony
with the laws of physics.

After a detailed introduction on the origins of ``logically
reversibile computer operations'' and an excursion into the origins of
``irreversible computer operations'', we will develop in detail three
reversible models of computation and discuss their potential
applications.

%%%%%%%%%
\paragraph*{Maxwell's Demon.}
To fully appreciate the missing principle of ``reversibility'' in
conventional computing, we go back to an old thought experiment by
J. C. Maxwell. The details are codified in a letter that Maxwell wrote
to P. G Tait in 1867 -- the letter, whose ideas are now known as
\emph{Maxwell's Demon}, tells of a thought experiment that seems to
indicate that intelligent beings can somehow violate the second law of
thermodynamics, thereby violating physics itself. Many resolutions
were offered for this conundrum (for a compilation, see the book by
Leff and Rex~\cite{leff1990}), but none withstood careful scrutiny
until the establishment of \emph{Landauer's Principle} in
1961~\cite{Landauer:1961} -- a principle whose experimental validation
happened recently in 2012~\cite{berut2012experimental}.

Maxwell's demon appears to violate the second law of thermodynamics by
having a tiny ``intelligence'' observing the movement of individual
particles of a gas and separating fast moving particles from slow
moving ones, thereby reducing the total entropy of the
system. Landauer's resolution of the demon relied on two ideas that
took root only a few decades earlier: the formal notion of computation
(through the work of Turing and Church, 1936) and the formal notion of
information (through the work of Shannon, 1948). Landauer reasoned
that the computation done by the finite brain of the demon, involves
getting information about the movement of molecules, storing that
information, analyzing that information to act on it, and then --- and
this is the critical step --- overwriting it to make room for the next
computation.  In other words, the computation that is manipulating
information in the demon's brain \textit{must be thermodynamic work},
thereby bringing the demon back into the fold of physics.

This is a strange and wonderful idea: information, physics, and
computation are inextricably linked. In contrast, when the early
models of computation were developed, there was no compelling reason
to the take the information content of computations into consideration
-- in fact, at that time there was no quantifiable notion of
information. These models followed in the footsteps of logic where,
following hundreds of years of tradition, the truth of a statement was
seen as ``absolute'' and independent of any reasoning, understanding,
or action. Statements were either true or false with no regard to any
``observer'' and the idea that statements had information content that
should be preserved was outside the classical understanding of
logic. Hence the fact that conventional logic operations such as
conjunction and disjunction were logically irreversible and hence lose
information was not a concern. Landauer's observation implied however
that ideas in each field have consequences for the
other~\cite{bennett:1973:lrc,bennett1985fundamental,bennett2010notes,bennett2003notes,baker:1992:nft,baez2011physics,dblp:conf/csfw/malacarias12}. To
really appreciate this fact, we delve deeper into the origin of our
computational models and argue that they are essentially reflections
of contemporary laws of physics.

%%%%%%%%%
\paragraph*{Origins of Computational Models.}
Current high-level programming languages as well as current hardware
are both based on Boolean Logic -- which really ought to be called
\emph{Piercean logic}, as Boole had a rather different conception of
logic which is rather far from our current models.  Nevertheless,
going back to Boole's 1853 book entitled \emph{An Investigation of the
  Laws of Thought, on which are Founded the Mathematical Theories of
  Logic and Probabilities}, we find that the opening sentence of Ch.~1 is:
\begin{quote}
  The design of the following treatise is to investigate the fundamental laws
  of those operations of the mind by which reasoning is performed; \ldots
\end{quote}
which clearly identifies the ``source'' of the logical laws as
mirroring Boole's understanding of human reasoning. A few chapters
later, we find:
\begin{quote}
  \textbf{Proposition IV.}  That axiom of metaphysicians which is termed the
  principle of contradiction, and which affirms that it is impossible for any
  being to possess a quality, and at the same time not to possess it, is a
  consequence of the fundamental law of thought, whose expression is $x^2 =
  x$.
\end{quote}
This ``law'' is reasonable in a classical world but is violated by the
postulates of quantum mechanics. Although a detailed historical
analysis of Boole's ideas in the light of modern physics is beyond our
scope, the above quotes should convey the idea that our elementary
computing notions date back to ideas that were thought reasonable in
the late 1800s.

Machines that ``compute'' are quite old. M\"{u}ller (1786) first
conceived of the idea of a ``difference machine'', which Babbage
(1819--1822) was able to construct. There are other computer
precursors as well -- the first stored programs were actually for
looms, most notably those of Bouchon (1725) which operated on a paper
tape, and Jacquard (1804) which operated by chains of punched cards.
But it was only in the 20th century that computer science emerged as a
formal discipline. One of the pioneering works was Alan Turing's
seminal paper in 1936 which established the idea that computation has
a formal interpretation and that all computability can be captured
within a formal system. Implicit in this achievement however is the
idea that abstract models of computation are just that --
\emph{abstractions of computation realized in the physical world.}
Indeed, going back to Turing's 1936 article \emph{On Computable
  Numbers, with an Application to the Entscheidungsproblem,}, the
opening sentence of Sec. 1 is:
\begin{quote}
  We have said that the computable numbers are those whose decimals are
  calculable by finite means \ldots the justification lies in the fact that
  the human memory is necessarily limited.
\end{quote}
In Sec. 9, we find:
\begin{quote}
  I think it is reasonable to suppose that they can only be squares
  whose distance from the closest of the immediately previously
  observed squares does not exceed a certain fixed amount.
\end{quote}
It is worth noting that these assumptions are both physical (on
distances) and metaphysical (on restrictions of the mind).  If we take
the human mind to be a physical ``machine'' which does computation,
then when both of the above assumptions are translated to the language
of physics, they embody what is known as the ``Bekenstein bound'',
which is an upper limit on the amount of information that can be
contained within a given finite region of space. A detailed historical
account of these ideas in the context of modern physics is again
beyond our scope. However, the quotes above, like the ones before,
should convey the ideas that our theories of computation and
complexity are based on some physical assumptions that Turing and
others found reasonable in the 1930s.

To summarize, a major achievement of computer science has been the
development of abstract models of computation that shield the
discipline from rapid changes in the underlying technology. Yet, as
effective as these models have been, one must note that they
\emph{embody several implicit physical assumptions} and these
assumptions are based on a certain understanding of the laws of
physics. Our understanding of physics has evolved tremendously since
1900!  Thus it is time to revisit these abstractions, especially with
respect to quantum mechanics.  Indeed one should take the physical
principles underlying quantum mechanics, the most successful physical
theory known to us and adapt computation to ``learn'' from these
principles.  In the words of Girard:
\begin{quote}
  In other terms, what is so good in logic that quantum physics should
  obey?  Can't we imagine that our conceptions about logic are wrong,
  so wrong that they are unable to cope with the quantum miracle?
  [\ldots] Instead of teaching logic to nature, it is more reasonable
  to learn from her. Instead of interpreting quantum into logic, we
  shall interpret logic into quantum (Girard 2007).
\end{quote}

@@

When we look around, we find many different words for related
concepts: isomorphism, equivalence, sameness, equality,
interchangeability, comparability, and correspondence, to name a
few. Some of these are informal concepts, while others have formal
mathematical meaning.  More importantly, even amongst the formal
concepts, there are differences -- which is why there are so many of
them! Because there are many such notions, we also need to walk our
way through them to find the one which is ``just right''. Thus we seek
a concept which is neither too strong nor too weak, that will express
when some structured information should be treated as ``the same''.
We can draw an analogy with topology: in topology, all point sets can
always be equipped with either the discrete or the indiscrete
topology, but both of these extremes are rarely useful. We will
develop our working notion of ``sameness'' as we go through the
various components that make up a programming language.

@@

This discussion is purely ``semantic'', in the sense that it is about
the denotation of simple primitive data (sets) and their reversible
deformations (bijections).  We would like to reverse engineer a programming
language from this denotation. But first, an obvious remark: any two sets
$C$ and $D$ of cardinality $n$ are always in bijective correspondence. So
we can abstract away from the details of the elements of $C$ and $D$ and
instead choose canonical representations -- in much the same way as computers
choose binary words to represent everything.

@@

So mere inhabitation falls far short of our goals of being able to
smoothly deform from one type to another. Let us thus analyze the crux
of the ``problem''. In logic, we have that $\land$ and $\lor$ are both
\emph{idempotent}: this is the property of any binary operation $*$
where $\forall a. a * a = a$. And it should be clear that an
idempotent operations is a \emph{forgetful} operation: its input has
two copies of $a$, but its output, only one. On the type side,
something more subtle happens. Consider $\top \uplus \top$
versus $\top$; the first has \emph{two} proofs of inhabitation
($\texttt{inl tt}$ and $\texttt{inr tt}$) while the second only one
($\texttt{tt}$). These cannot be put in bijective correspondence. Even
though the ``payload'' \texttt{tt} is the same, forgetting
$\texttt{inl}$ (or \texttt{inr}) throws away information -- something
we have expressly disallowed.  Yes, this should remind you of
Maxwell's demon: even though the data is the same, they are tagged
differently, and this data is indeed information, and must be
preserved.

@@

\amr{***HERE***}

@@

Thus, rather than looking at (untyped) commutative semirings,
we should looked at a \emph{typed} version. This process frequently
goes by the moniker of ``categorification''.  We want a categorical
algebra, where the basic objects are groupoids (to model our types),
and where there is a natural notion of $+$ and $*$.  At first,
we hit what seems like a serious stumbling block: the category of
all groupoids, Groupoid, does not have co-products or products.
We don't want to work internally in Groupoid -- we want operations
\emph{on} groupoids. In other words, we want something akin to
symmetric monoidal categories, but with two interacting 
monoidal structures.  Luckily, this already exists: the categorical
analog to (commutative) semirings are (symmetric) Rig Categories.
This straightforwardly generalizes to symmetric Rig Groupoids.

@@

DANGER: here, we leave the safe land of ``preservation of information'',
because partiality is not reversible. So while trace gives a reversible
program as output, it itself isn't a reversible operation. Like the
demon's brain in Maxwell's thought experiment, trace irreversible work!
So trace, which is fundamental to all work with recursive types, is
only reversible "at level 1". So we need to replace trace with some other
operation for level 2 reversibility to also hold -- and this is an open
problem.

@@

\amr{We should also have a quick survey of other works in reversible
computation? Most of the other work tries to start from well known
models of computation or well known programming languages, and then
adapts them to be reversible. Pi is different in that it starts from
the semantic implications of reversibility. Then, because it adds
types, rather than control-flow (or even data-flow) as its next layer
of ``understanding'', this leads to equivalences, isomorphisms, etc.
This then leads, quite naturally, to finding that the ``proof
language'' of semirings (and Rig Groupoids at level 2) is actually a
programming language. And it is Pi. This is a neat twist on
Curry-Howard because CH is about \textbf{inhabitation} only. But with
``conservation of information'' as the basis, a different kind of
correspondence arises; in fact, this one may well be an actual
isomorphism.}

%%%%%%%%%
\paragraph*{Information.}
We can now give a first precise definition of information preservation.
Let
`$b$' be a (not necessarily finite) type whose values are labeled
$b^1, b^2, \ldots$. Let $\xi$ be a random variable of type $b$ that is
equal to $b^i$ with probability $p_i$. The entropy of $\xi$ is defined
as $-\sum p_i \log{p_i}$~\cite{Shannon1948}.  Consider a function
$f : b_1 \rightarrow b_2$ where $b_2$ is a (not necessarily finite)
type whose values are labeled $b_2^1, b_2^2, \ldots$. The output
entropy of the function is given by $- \sum q_j \log{q_j}$ where $q_j$
indicates the probability of the output of the function to have value
${b_2}^j$. We say a function is \emph{information-preserving} if its
output entropy is equal to the entropy of its input. In later sections,
we will refine this view.

@@

\bibliographystyle{acm}
%\pagestyle{headings}
\bibliography{cites}
\end{document}

>>>>

\noindent These quotes by Toffoli and Feynman both highlight the
consequences of two obvious observations: (i) all the operations that
a computer performs reduce to basic physical operations; and (ii)
there is a mismatch between the logical operations of a typical model
of computation (which are logically irreversible) and the fundamental
laws of physics (which are reversible). One could certainly dismiss
the mismatch as irrelevant to the practice of computing but our thesis
is that the next computing revolution is likely to be founded on
revised models of computation that are designed to be in closer harmony
with the laws of physics.

After a detailed introduction on the origins of ``logically
reversibile computer operations'' and an excursion into the origins of
``irreversible computer operations,'' we will develop in detail three
reversible models of computation and discuss their potential
applications.

%%%%%%%%%
\paragraph*{Maxwell's Daemon.}
To fully appreciate the missing principle of ``reversibility'' in
conventional computing, we go back to an old thought experiment by
J. C. Maxwell. The details are codified in a letter that Maxwell wrote
to P. G Tait in 1867 -- the letter, whose ideas are now known as
\emph{Maxwell's Daemon}, tells of a thought experiment that seems to
indicate that intelligent beings can somehow violate the second law of
thermodynamics, thereby violating physics itself. Many resolutions
were offered for this conundrum (for a compilation, see the book by
Leff and Rex~\cite{leff1990}), but none withstood careful scrutiny
until the establishment of \emph{Landauer's Principle} in
1961~\cite{Landauer:1961} -- a principle whose experimental validation
happened recently in 2012~\cite{berut2012experimental}.

Maxwell's daemon appears to violate the second law of thermodynamics by
having a tiny ``intelligence'' observing the movement of individual
particles of a gas and separating fast moving particles from slow
moving ones, thereby reducing the total entropy of the
system. Landauer's resolution of the daemon relied on two ideas that
took root only a few decades earlier: the formal notion of computation
(through the work of Turing~\cite{turing}, Church~\cite{church51}, and
others) and the formal notion of information (through the work of
Shannon~\cite{shannon1948}). Landauer reasoned that the computation
done by the finite brain of the daemon, involves getting information
about the movement of molecules, storing that information, analyzing
that information to act on it, and then --- and this is the critical
step --- overwriting it to make room for the next computation.  In
other words, the computation that is manipulating information in the
daemon's brain \textit{must be thermodynamic work}, thereby bringing
the daemon back into the fold of physics.

This is a strange and wonderful idea: information, physics, and
computation are inextricably linked. In contrast, when the early
models of computation were developed, there was no compelling reason
to take the information content of computations into consideration
-- in fact, at that time there was no quantifiable notion of
information. These models followed in the footsteps of logic where,
following hundreds of years of tradition, the truth of a statement was
seen as ``absolute'' and independent of any reasoning, understanding,
or action. Statements were either true or false with no regard to any
``observer'' and the idea that statements had information content that
should be preserved was outside the classical understanding of
logic. Hence the fact that conventional logic operations such as
conjunction and disjunction were logically irreversible and hence lose
information was not a concern. Landauer's observation implied however
that ideas in each field have consequences for the
other~\cite{bennett:1973:lrc,bennett1985fundamental,bennett2010notes,bennett2003notes,baker:1992:nft,baez2011physics,dblp:conf/csfw/malacarias12}. To
really appreciate this fact, we delve deeper into the origin of our
computational models and argue that they are essentially reflections
of contemporary laws of physics.

%%%%%%%%%
\paragraph*{Origins of Computational Models.}
Current high-level programming languages as well as current hardware
are both based on the mathematical formalization of logic developed by
De Morgan, Venn, Boole, and Peirce in the mid to late 1800s. Going
back to Boole's 1853 book entitled \emph{An Investigation of the Laws
  of Thought, on which are Founded the Mathematical Theories of Logic
  and Probabilities}, we find that the opening sentence of Ch.~1 is:
\begin{quote}
  The design of the following treatise is to investigate the fundamental laws
  of those operations of the mind by which reasoning is performed; 
\end{quote}
which clearly identifies the ``source'' of the logical laws as
mirroring Boole's understanding of human reasoning. A few chapters
later, we find:
\begin{quote}
  \textbf{Proposition IV.}  That axiom of metaphysicians which is termed the
  principle of contradiction, and which affirms that it is impossible for any
  being to possess a quality, and at the same time not to possess it, is a
  consequence of the fundamental law of thought, whose expression is $x^2 =
  x$.
\end{quote}
This ``law'' is reasonable in a classical world but is violated by the
postulates of quantum mechanics. Although a detailed historical
analysis of Boole's ideas in the light of modern physics is beyond our
scope, the above quotes should convey the idea that our elementary
computing notions date back to ideas that were thought reasonable in
the late 1800s.

Machines that ``compute'' are quite old. M\"{u}ller (1786) first
conceived of the idea of a ``difference machine,'' which Babbage
(1819--1822) was able to construct. There are other computer
precursors as well -- the first stored programs were actually for
looms, most notably those of Bouchon (1725) which operated on a paper
tape, and Jacquard (1804) which operated by chains of punched cards.
But it was only in the 20th century that computer science emerged as a
formal discipline. One of the pioneering works was Alan Turing's
seminal paper~\cite{turing} in 1936 which established the idea that
computation has a formal interpretation and that all computability can
be captured within a formal system. Implicit in this achievement
however is the idea that abstract models of computation are just that
-- \emph{abstractions of computation realized in the physical world.}
Indeed, going back to Turing's 1936 article \emph{On Computable
  Numbers, with an Application to the Entscheidungsproblem,}, the
opening sentence of Sec. 1 is:
\begin{quote}
  We have said that the computable numbers are those whose decimals
  are calculable by finite means [\ldots] the justification lies in
  the fact that the human memory is necessarily limited.
\end{quote}
In Sec. 9, we find:
\begin{quote}
  I think it is reasonable to suppose that they can only be squares
  whose distance from the closest of the immediately previously
  observed squares does not exceed a certain fixed amount.
\end{quote}
It is worth noting that these assumptions are both physical (on
distances) and metaphysical (on restrictions of the mind).  If we take
the human mind to be a physical ``machine'' which does computation,
then when both of the above assumptions are translated to the language
of physics, they embody what is known as the ``Bekenstein
bound''~\cite{PhysRevD.23.287}, which is an upper limit on the amount
of information that can be contained within a given finite region of
space. A detailed historical account of these ideas in the context of
modern physics is again beyond our scope. However, the quotes above,
like the ones before, should convey the ideas that our theories of
computation and complexity are based on some physical assumptions that
Turing and others found reasonable in the 1930s.

To summarize, a major achievement of computer science has been the
development of abstract models of computation that shield the
discipline from rapid changes in the underlying technology. Yet, as
effective as these models have been, one must note that they
\emph{embody several implicit physical assumptions} and these
assumptions are based on a certain understanding of the laws of
physics. Our understanding of physics has evolved tremendously since
1900!  Thus it is time to revisit these abstractions, especially with
respect to quantum mechanics.  Indeed one should take the physical
principles underlying quantum mechanics, the most successful physical
theory known to us and adapt computation to ``learn'' from these
principles.  In the words of Girard~\cite{Girard:2007:TMI:1348911.1348915}:
\begin{quote}
  In other terms, what is so good in logic that quantum physics should
  obey?  Can't we imagine that our conceptions about logic are wrong,
  so wrong that they are unable to cope with the quantum miracle?
  [\ldots] Instead of teaching logic to nature, it is more reasonable
  to learn from her. Instead of interpreting quantum into logic, we
  shall interpret logic into quantum.
\end{quote}

@@

When we look around, we find many different words for related
concepts: isomorphism, equivalence, sameness, equality,
interchangeability, comparability, and correspondence, to name a
few. Some of these are informal concepts, while others have formal
mathematical meaning.  More importantly, even amongst the formal
concepts, there are differences -- which is why there are so many of
them! Because there are many such notions, we also need to walk our
way through them to find the one which is ``just right.'' Thus we seek
a concept which is neither too strong nor too weak, that will express
when some structured information should be treated as ``the same.''
We can draw an analogy with topology: in topology, all point sets can
always be equipped with either the discrete or the indiscrete
topology, but both of these extremes are rarely useful. We will
develop our working notion of ``sameness'' as we go through the
various components that make up a programming language.

@@

This discussion is purely ``semantic,'' in the sense that it is about
the denotation of simple primitive data (sets) and their reversible
deformations (bijections).  We would like to reverse engineer a programming
language from this denotation. But first, an obvious remark: any two sets
$C$ and $D$ of cardinality $n$ are always in bijective correspondence. So
we can abstract away from the details of the elements of $C$ and $D$ and
instead choose canonical representations -- in much the same way as computers
choose binary words to represent everything.

@@

So mere inhabitation falls far short of our goals of being able to
smoothly deform from one type to another. Let us thus analyze the crux
of the ``problem.'' In logic, we have that $\land$ and $\lor$ are both
\emph{idempotent}: this is the property of any binary operation $*$
where $\forall a. a * a = a$. And it should be clear that an
idempotent operations is a \emph{forgetful} operation: its input has
two copies of $a$, but its output, only one. On the type side,
something more subtle happens. Consider $\top \uplus \top$
versus $\top$; the first has \emph{two} proofs of inhabitation
($\texttt{inl tt}$ and $\texttt{inr tt}$) while the second only one
($\texttt{tt}$). These cannot be put in bijective correspondence. Even
though the ``payload'' \texttt{tt} is the same, forgetting
$\texttt{inl}$ (or \texttt{inr}) throws away information -- something
we have expressly disallowed.  Yes, this should remind you of
Maxwell's daemon: even though the data is the same, they are tagged
differently, and this data is indeed information, and must be
preserved.

@@

\amr{***HERE***; operational semantics; It feels like I should be
  writing about how the combinators (both for spaces and for
  'permutations') work at a lower level. Did we ever write that up
  somewhere?  We have lots of Agda code which 'does it', nicely, but
  somehow I can't remember an English prose version of it. I am not
  sure how long we want to make it.  For example, do we talk about the
  'matrix' representation of combinators-for-permutations-on-[n], and
  how these make things like 0+a=a and associativity obvious by
  construction?
  Let’s have a geometric representation of spaces:
- A + B means putting A next to B
- A * B means putting A and B on two perpendicular axes and filling the square (volume) between them
All the combinators would indeed have a natural action and if we have
a few nice pictures it would be great. What do you think? }

@@

Thus, rather than looking at (untyped) commutative semirings,
we should looked at a \emph{typed} version. This process frequently
goes by the moniker of ``categorification.''  We want a categorical
algebra, where the basic objects are groupoids (to model our types),
and where there is a natural notion of $+$ and $*$.  At first,
we hit what seems like a serious stumbling block: the category of
all groupoids, Groupoid, does not have co-products or products.
We don't want to work internally in Groupoid -- we want operations
\emph{on} groupoids. In other words, we want something akin to
symmetric monoidal categories, but with two interacting 
monoidal structures.  Luckily, this already exists: the categorical
analog to (commutative) semirings are (symmetric) Rig Categories.
This straightforwardly generalizes to symmetric Rig Groupoids.

@@

DANGER: here, we leave the safe land of ``preservation of information,''
because partiality is not reversible. So while trace gives a reversible
program as output, it itself isn't a reversible operation. Like the
daemon's brain in Maxwell's thought experiment, trace irreversible work!
So trace, which is fundamental to all work with recursive types, is
only reversible "at level 1". So we need to replace trace with some other
operation for level 2 reversibility to also hold -- and this is an open
problem.

@@

\amr{We should also have a quick survey of other works in reversible
computation? Most of the other work tries to start from well known
models of computation or well known programming languages, and then
adapts them to be reversible. Pi is different in that it starts from
the semantic implications of reversibility. Then, because it adds
types, rather than control-flow (or even data-flow) as its next layer
of ``understanding,'' this leads to equivalences, isomorphisms, etc.
This then leads, quite naturally, to finding that the ``proof
language'' of semirings (and Rig Groupoids at level 2) is actually a
programming language. And it is Pi. This is a neat twist on
Curry-Howard because CH is about \textbf{inhabitation} only. But with
``conservation of information'' as the basis, a different kind of
correspondence arises; in fact, this one may well be an actual
isomorphism.}

%%%%%%%%%
\paragraph*{Information.}
We can now give a first precise definition of information preservation.
Let
`$b$' be a (not necessarily finite) type whose values are labeled
$b^1, b^2, \ldots$. Let $\xi$ be a random variable of type $b$ that is
equal to $b^i$ with probability $p_i$. The entropy of $\xi$ is defined
as $-\sum p_i \log{p_i}$~\cite{shannon1948}.  Consider a function
$f : b_1 \rightarrow b_2$ where $b_2$ is a (not necessarily finite)
type whose values are labeled $b_2^1, b_2^2, \ldots$. The output
entropy of the function is given by $- \sum q_j \log{q_j}$ where $q_j$
indicates the probability of the output of the function to have value
${b_2}^j$. We say a function is \emph{information-preserving} if its
output entropy is equal to the entropy of its input. In later sections,
we will refine this view.

@@

\bibliographystyle{unsrt}
%\pagestyle{headings}
\bibliography{cites}
\end{document}
